---
title: "MA 4710 Homework 9"
author: "Benjamin Hendrick"
date: "April 6, 2016"
output: 
  pdf_document:
    highlight: null
---

# Problem 7.3 

Load the data and rename the columns.

```{r}
brand <- read.table("~/GitHub/MA-4710/Homework 8/brand.txt", quote="\"", comment.char="")
names(brand) <- c("Yi","Xi1", "Xi2")
```

## Part A

Find the linear models for $Y+i = \beta_1 X_1$, $Y_i = \beta_2 X_2$, and $Y_i = \beta_1 X_1 + \beta_2 X_2$

```{r}
brand.fit1 <- lm(Yi ~ Xi1, data = brand)
brand.fit2 <- lm(Yi ~ Xi2, data = brand)
brand.fit3 <- lm(Yi ~ Xi1 + Xi2, data = brand)
```

Find the ANOVA table for the third model ($Y_i = \beta_1 X_1 + \beta_2 X_2$) using the `anova` function.

```{r}
anova(brand.fit3) 
```

From the above outpout, the following ANOVA table is produced:

Name | $df$ | $SS$ | $MS$ | $F^*$
-----|------|------|------|-------
Responce | 2 | 1872.8 | 936.4 | 215.947
$X_{1}$ | 1 | 1566.45 | 1566.45 | 216.06
$X_{2}$ | 1 | 306.25 | 306.25 | 42.219
Error | 13 | 94.30 | 7.25 | 
Total | 17 | 1967.1 | |

## Part B

$$H_0 : \beta_2 = 0$$
$$H_a : \beta_2 \ne 0$$

```{r}
f <- (anova(brand.fit3)[2,2]/anova(brand.fit3)[2,1])/
               ((anova(brand.fit3)[2,2]+anova(brand.fit3)[3,2])/
                  (anova(brand.fit3)[2,1]+anova(brand.fit3)[3,1]))
```

The F test statistic to see if $X_{2}$ can be dropped is `r f`. The critical value is `r df(0.01, anova(brand.fit3)[2,1],(anova(brand.fit3)[2,1]+anova(brand.fit3)[3,1]))`

Becasue the test statistic is greater than the critical value, we reject the null hypothesis $H_{0}$. Therefore we should not drop $X_{2}$.

# Problem 7.12

```{r}
R2_Y1 <- anova(brand.fit1)[1,2]/(anova(brand.fit1)[1,2]+anova(brand.fit1)[2,2])
R2_Y2 <- anova(brand.fit2)[1,2]/(anova(brand.fit2)[1,2]+anova(brand.fit2)[2,2])
R2_12 <- (anova(brand.fit3)[1,2]+anova(brand.fit3)[2,2])/
          (anova(brand.fit3)[1,2]+anova(brand.fit3)[2,2]+anova(brand.fit3)[3,2])
R2_Y12 <- anova(brand.fit3)[1,2]/(anova(brand.fit3)[1,2]+anova(brand.fit3)[3,2])
R2_Y21 <- anova(brand.fit3)[2,2]/(anova(brand.fit3)[2,2]+anova(brand.fit3)[3,3])
R2 <- (anova(brand.fit3)[1,2]+anova(brand.fit3)[2,2])/
  (anova(brand.fit3)[1,2]+anova(brand.fit3)[2,2]+anova(brand.fit3)[3,2])
```

$R^2_{Y1} =$ `r R2_Y1` percent of the error in the total model can be explained by $X_1$.

$R^2_{Y2} =$ `r R2_Y2` percent of the error in the total model can be explained by $X_2$.

$R^2_{12} =$ `r R2_12` percent of the error in the total model can be explained by $X_1$ and $X_2$. 

$R^2_{Y1|2} =$ `r R2_Y12` percent of the error in $X_2$ can be explained by $X_1$.

$R^2_{Y2|1} =$ `r R2_Y21` percent of the error in $X_1$ can be explained by $X_2$.

$R^2 =$ `r R2` percent of the error in the total model can be explained by $X_1$ and $X_2$.

# Problem 7.16

## Part A
The following code transforms the data and standardizes the regression model. 

```{r}
n <- dim(brand)[1]
brand.trans <- data.frame(1/sqrt(n-1)*scale(brand))
names(brand.trans) <- c("tYi", "tXi1", "tXi2")
brand.trans.fit <- lm(tYi ~ -1 + tXi1 + tXi2, data = brand.trans) 
```

## Part B

```{r}
brand.trans.fit$coefficients 
```

The transformed coefficient $^*b_1$ means that the transformed $X_1$ increased 0.8924 per standard deviation.

## Part C


# Problem 7.24

## Part A
```{r}
brand.simple <- lm(Yi ~ Xi1, data = brand)
brand.simple$coefficients
```

The simple regression model is $Y_{i} = 50.775 + 4.425 X_1$.

## Part B
```{r}
brand.lm.6.5 <- lm(Yi ~ Xi1 + Xi2, data=brand)
brand.lm.6.5$coefficients
```

The coefficients for moisture content are the same between Problem 7.24a and Problem 6.5b.

## Part C
```{r}
anova(brand.simple)[1,2]
anova(brand.lm.6.5)[1,2]
```

$SSR(X_1) =$ `r anova(brand.simple)[1,2]` and $SSR(X_1 | X_2) =$ `r anova(brand.lm.6.5)[1,2]` are equal.

## Part D
Recall the correlation matrix from Problem 6.5a.

```{r}
cor(brand)
```

According to the matrix, there is no correlation between $X_1$ and $X_2$. Therefore, Part B is justified because $X_2$ has no influence on $X_1$. Likewise, Part C is justified because the error of $X_2$ has no influence on $X_1$.
