---
title: "MA 4710 Homework 2"
author: "Benjamin Hendrick"
date: "February 8, 2016"
output: pdf_document
---

# Problem 2.13
Load the data.

```{r}
CH01PR19 <- read.table("/var/folders/6j/_n_8wtg912559sqd775wfctw0000gn/T//Rtmp8EbqFk/datacdf786779fb", 
                       quote="\"", comment.char="")
```

## Part A

Obtain the 95% confidence interval using the `confint` function in R.

```{r}
model <- lm(V1 ~ V2, data = CH01PR19)
confint(model)
```

Based on the output above, the 95% confidence interval is (0.01353307, 0.06412118). Because zero is not in the confidence interval, we can assume that slope is not zero.

## Part B

```{r}
predict(model, CH01PR19, interval="predict")[120,]
```

The prediction interval for a score of 28 is (1.959355, 4.443063). The prediction interval contains the fitted score (3.201209). This means that we can be 95% sure that the predicted GPA, with an ACT score of 28, will be between 1.959355 and 4.443063.

## Part C

The prediction interval is larger than the confidence interval. The prediction interval should be larger because it takes into account the variation between $Y$ and $X$.

## Part D
```{r}
yh <- 2.11405 + (28)*0.03883
w2 <- qf(.95,2,118)
yh + sqrt(w2)*yh
yh - sqrt(w2)*yh
```

The boundary values of the 95% confidence interval are (-2.410646, 8.813226). This interval is larger than the interval in part B. This interval shouldn't be much larger than the prediction interval.


# Problem 2.23

Use the same data and model as in Problem 2.13

## Part A

```{r}
anova(model)
```

The code above provides the following ANOVA table.

|       | $df$ | SS     | MS     | F      |
|-------|------|--------|--------|--------|
| Model | 1    | 3.588  | 3.588  | 9.2402 |
| Error | 118  | 45.818 | 0.3883 |        |
| Total | 119  | 49.507 |        |        |

## Part B

MSE estimates the variance. MSR estimates the variance plus the residual error. MSR and MSE predict the same values if the residual error is zero.

## Part C

State the hypothesis:

\begin{align*}
H_{0} &: \beta_{0} = 0 \\
H_{1} &: \beta_{0} \ne 0
\end{align*}

Reject the null hypothsis if the F value from the ANOVA table is greater than the critical value. Find the critical value.

$$F_{0.01}(1,118) = 6.854641$$

We reject the null hypothesis becasue the test statistic (9.2402) is greater than the critical value (6.854641).

# Problem 2.26

Read the data.

```{r}
CH01PR22 <- read.table("/var/folders/6j/_n_8wtg912559sqd775wfctw0000gn/T//Rtmp8EbqFk/datacdf58d91d97", 
                       quote="\"", comment.char="")
```

## Part A

```{r}
model <- lm(V1~V2, data = CH01PR22)
anova(model)
```

The code above provides the following ANOVA table.

|       | $df$ | SS      | MS     | F      |
|-------|------|---------|--------|--------|
| Model | 1    | 5297.5  | 5297.5 | 506.51 |
| Error | 14   | 146.4   | 10.5   |        |
| Total | 15   | 5,443.9 |        |        |

## Part B

State the hypothesis:

\begin{align*}
H_{0} &: \beta_{0} = 0 \\
H_{1} &: \beta_{0} \ne 0
\end{align*}

Reject the null hypothsis if the F value from the ANOVA table is greater than the critical value. Find the critical value.

$$F_{0.01}(1,14) = 8.861593$$

We reject the null hypothesis becasue the test statistic (506.51) is greater than the critical value (8.861593).

## Part C

```{r}
plot(model)
```

The two plots of interest are the residual plot (first plot) and the QQ plot (second plot). Based on the plots, the SSR appears to be a larger component of the SSTO. This implies that the magnitude of $R^{2}$ will be large.

## Part D

```{r}
summary(model)
```

From the code above, $R^{2} = 0.9731$ and $r = \sqrt{0.9731} = 0.9865$.