---
title: "MA 4710 Homework 6"
author: "Benjamin Hendrick"
date: "March 1, 2016"
output: pdf_document
---

# Problem 4.5

Load the data into `R` and rename the variables.

```{r}
filePath <- "~/GitHub/MA-4710/Homework 6/CH01PR22.txt"
CH01PR22 <- read.table(filePath, quote="\"", comment.char="")
names(CH01PR22)[1] <- "hardness"
names(CH01PR22)[2] <- "time"
```

## Part A

Create the linear model using the `lm` function and find the coefficients with the `summary` function.

```{r}
plastic.lm <- lm(hardness ~ time, data = CH01PR22)
plastic.coef <- summary(plastic.lm)$coefficients         
alpha <- 0.1
```

Compute the Bonferroni join confidence intervals for $\beta_{0}$ and $\beta_{1}$, using a 90 percent family confidence coefficient.

```{r}
B <- qt(1-alpha/(2*2),plastic.lm$df.residual) 
BCI <- cbind(plastic.coef[,1]-B*plastic.coef[,2],
             plastic.coef[,1]+B*plastic.coef[,2]) 
colnames(BCI) <- c("Lower Bound","Upper Bound") 
```

The confidence interval for the intercept $\beta_{0}$ is (`r BCI[1,1]`, `r BCI[1,2]`). We conclude that the intercept $\beta_{0}$ has a 90 percent chance of being between `r BCI[1,1]` and `r BCI[1,2]`.

The confidence interval for the slope $\beta_{1}$ is (`r BCI[2,1]`, `r BCI[2,2]`).  We conclude that the slope $\beta_{1}$ has a 90 percent chance of being between `r BCI[2,1]` and `r BCI[2,2]`.


# Problem 4.9

## Part A

Using the data from Problem 4.5, create the 90 percent confidence interval Bonferroni bands at 20, 30, and 40 hours.

```{r}
CI <- predict(plastic.lm,newdata=data.frame(time=c(20,30,40)),se.fit=TRUE) 
g <- 3
B <- qt(1-alpha/(2*g),plastic.lm$df.residual) 
BBand <- cbind( CI$fit - B * CI$se.fit, CI$fit + B * CI$se.fit )
BBand
```

The confidence interval at 20 hours is (`r BBand[1,1]`, `r BBand[1,2]`). We conclude that at 20 hours, the estimate has a 90 chance of being between `r BBand[1,1]` and `r BBand[1,2]`.

The confidence interval at 30 hours is (`r BBand[2,1]`, `r BBand[2,2]`). We conclude that at 30 hours, the estimate has a 90 chance of being between `r BBand[2,1]` and `r BBand[2,2]`.

The confidence interval at 40 hours is (`r BBand[3,1]`, `r BBand[3,2]`). We conclude that at 40 hours, the estimate has a 90 chance of being between `r BBand[3,1]` and `r BBand[3,2]`.

## Part C

Create the the 90 percent confidence interval Working-Hotelling bands at 30 and 40.

```{r}
CI <- predict(plastic.lm,newdata=data.frame(time=c(30,40)),se.fit=TRUE)
              
W <- sqrt(2*qf(0.90,length(plastic.lm$coefficients),plastic.lm$df.residual))
WHBand <- cbind( CI$fit - W * CI$se.fit, CI$fit + W * CI$se.fit )
```

The confidence interval at 30 hours is (`r WHBand[1,1]`, `r WHBand[1,2]`). We conclude that at 30 hours, the estimate has a 90 chance of being between `r WHBand[1,1]` and `r WHBand[1,2]`.

The confidence interval at 40 hours is (`r WHBand[2,1]`, `r WHBand[2,2]`). We conclude that at 40 hours, the estimate has a 90 chance of being between `r WHBand[2,1]` and `r WHBand[2,2]`.

Create the the 90 percent confidence interval Bonferroni bands at 30 and 40.

```{r}
g <- 2
B <- qt(1-alpha/(2*g),plastic.lm$df.residual) 
BBand <- cbind( CI$fit - B * CI$se.fit, CI$fit + B * CI$se.fit )
BBand
```

The confidence interval at 30 hours is (`r BBand[1,1]`, `r BBand[1,2]`). We conclude that at 30 hours, the estimate has a 90 chance of being between `r BBand[1,1]` and `r BBand[1,2]`.

The confidence interval at 40 hours is (`r BBand[2,1]`, `r BBand[2,2]`). We conclude that at 40 hours, the estimate has a 90 chance of being between `r BBand[2,1]` and `r BBand[2,2]`.

Based on the above confidence intervals, the Bonferroni model is the most efficient because it has the smallest confience bands.

# Problem 5.1

Define the matrices $A$, $B$, and $C$ in `R`. 

```{r}
A <- matrix(c(1,2,3,4,6,8),3,2)
B <- matrix(c(1,1,2,3,4,5),3,2)
C <- matrix(c(3,5,8,4,1,0),2,3)
```

Perform the matrix arithmetic in `R` to compute $A+B$, $A-B$, $AC$, $AB'$, and $B'A$.

Note that the function `t(A)` in `R` is equivalent to $A^{T}$ and `A %*% B` is equivalent to $AB$. 

```{r}
A+B #A+B
A-B #A-B
A %*% C #AC
A %*% t(B) #AB'
t(B) %*% A #B'A
```


# Problem 5.5

Load the data into `R` and rename the variables. 

```{r}
filePath <- "~/GitHub/MA-4710/Homework 6/CH05PR05.txt"
CH05PR05 <- read.table(filePath, quote="\"", comment.char="")
names(CH05PR05)[1] <- "city"
names(CH05PR05)[2] <- "loans"
```

Put the $X$ and $Y$ values into their respective matrices. $\textbf{Y}$ is a $6\times 1$ matrix. $\textbf{X}$ is a $6\times 2$ matrix. The first column of $\textbf{X}$ is all 1's and the second column is the values of $X$.

```{r}
Y <- matrix(CH05PR05$loans, 6,1)
X <- matrix(c(1,1,1,1,1,1,CH05PR05$city), 6,  2)
```

Perform the matrix arithmetic in `R` to compute $Y'Y$, $XX'$, and $X'Y$.

```{r}
t(Y) %*% Y #Y'Y
X %*% t(X) #XX'
t(X) %*% Y #X'Y
```

# Problem 5.13

Use the same data from Problem 5.5: `CH05PR05`.

Perform the matrix arithmetic in `R` to compute $(X'X)^{-1}$.

```{r}
(t(X) %*% X)^(-1) #(X'X)^-1
```